---
title: 'Comparing Business Regulation for Domestic Firms in 190 Economies: Equal Opportunity
  for All?'
author: "Shambhavi Mohan"
date: 'Due: October 27, 2017'
output:
  html_document: default
  pdf_document: default
urlcolor: green
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Background
The opportunity to find a job or ease of doing business can help people come out of poverty or prevent them from going in. I want to explore how the ease of doing business has improved across economies and how different lower income economies are as compared to high income ones.

*Important Definition:* An economy’s distance to frontier is reflected on a scale from 0 to 100, where 0 represents the lowest performance and 100 represents the frontier. For example, a score of 75 in DB 2016 means an economy was 25 percentage points away from the frontier constructed from the best performances across all economies and across time.


```{r, echo=FALSE, warning=FALSE, message=FALSE}

setwd("~/Documents/dataviz project")

#import libraries
library(gpclib)
library(rgeos)
library(sp) ## spatial data in R
library(rgdal) ## r bindings to geospatial data abstraction library
library(maptools)
library(broom)
library(dplyr)
library(ggplot2)
library(ggmap)
library(readr)
library(haven)
library(stringr)
library(tidyr)
library(RColorBrewer)
library(extrafont)
library(directlabels)
library(readr)
library(haven)
library(reshape2)
library(RColorBrewer)
library(extrafont)
library(directlabels)


doingbiz <- read.csv("doingbusiness.csv")
codes <- read.csv("UNSD — Methodology.csv")

#to find different names
#doingbiz[!(doingbiz$Economy %in% codes$Country.or.Area),]$Economy
#codes[!(codes$Country.or.Area%in% doingbiz$Economy),]$Country.or.Area

#Update the differences in the country names
doingbiz$Economy <- str_replace(doingbiz$Economy, "Bahamas, The" , "Bahamas")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Congo, Dem Rep" , "Democratic Republic of the Congo")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Congo, Rep", "Congo")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Cote d’Ivoire", "Côte d'Ivoire")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Egypt, Arab Rep", "Egypt")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Gambia, The", "Gambia")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Iran, Islamic Rep", 'Iran (Islamic Republic of)')
doingbiz$Economy <- str_replace(doingbiz$Economy, "Korea, Rep", "Republic of Korea")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Kyrgyz Republic", "Kyrgyzstan")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Lao PDR", "Lao People's Democratic Republic")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Macedonia, FYR", "Macedonia")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Micronesia, Fed Sts", "Micronesia (Federated States of)")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Puerto Rico US", "Puerto Rico")
doingbiz$Economy <- str_replace(doingbiz$Economy, "São Tomé and Príncipe", "Sao Tome and Principe")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Slovak Republic", "Slovakia")
doingbiz$Economy <- str_replace(doingbiz$Economy, "St ", "Saint ")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Tanzania", "United Republic of Tanzania")
doingbiz$Economy <- str_replace(doingbiz$Economy, "United Kingdom", "United Kingdom of Great Britain and Northern Ireland")
doingbiz$Economy <- str_replace(doingbiz$Economy, "United States", "United States of America")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Venezuela, RB", "Venezuela (Bolivarian Republic of)")
doingbiz$Economy <- str_replace(doingbiz$Economy, "West Bank and Gaza", "State of Palestine")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Yemen, Rep", "Yemen")
#doingbiz$Economy <- str_replace(doingbiz$Economy, "" , "")
codes$Country.or.Area <- str_replace(codes$Country.or.Area, "Bolivia (Plurinational State of)", "Bolivia")
codes$Country.or.Area <- str_replace(codes$Country.or.Area, "Czechia", "Czech Republic")
codes$Country.or.Area <- str_replace(codes$Country.or.Area, "The former Yugoslav Republic of Macedonia", "Macedonia")
codes$Country.or.Area <- str_replace(codes$Country.or.Area, "Republic of Moldova", "Moldova")
codes$Country.or.Area <- str_replace(codes$Country.or.Area, "Viet Nam", "Vietnam")
codes$Country.or.Area <- str_replace(codes$Country.or.Area, "Guam", "Micronesia (Federated States of)")
codes$Country.or.Area <- str_replace(codes$Country.or.Area, "China, Hong Kong Special Administrative Region", "Hong Kong SAR, China")
#codes$Country.or.Area <- str_replace(codes$Country.or.Area, "", "")
#Removing city names
doingbiz <- doingbiz[(!(doingbiz$Economy %in% c("Bangladesh - Chittagong", "Bangladesh - Dhaka", "Bolivia", "Brazil - Rio de Janeiro", "Brazil - São Paulo", "China - Beijing", "China - Shanghai", "India - Delhi", "India - Mumbai", "Indonesia - Jakarta", "Indonesia - Surabaya", "Japan - Osaka", "Japan - Tokyo", "Kosovo",  "Mexico - Mexico City","Mexico - Monterrey",  "Nigeria - Kano", "Nigeria - Lagos", "Pakistan - Karachi",  "Pakistan - Lahore",   "Russian Federation - Moscow", "Russian Federation - Saint Petersburg", "Taiwan, China", "United States of America - Los Angeles", "United States of America - New York City"))),]
#Merging codes and doingbiz
doingbiz <- merge(doingbiz, select(codes, c("Country.or.Area", "ISO.alpha3.Code")), by.x = "Economy", by.y = "Country.or.Area", all = FALSE)

doingbiz$Year <- gsub("DB*", "", doingbiz$Year) 
doingbiz$Year <- as.numeric(doingbiz$Year)
#doingbiz <- doingbiz %>% filter(!str_detect(Economy, "-"))

#Only distance to frontier data
frontier <- doingbiz %>% group_by(region,incomegroup, Year) %>% 
  summarise(StartingBusiness = mean(Starting.a.Business...DTF, rm.na = TRUE), 
  GettingConstructionPermits = mean(Dealing.with.Construction.Permits...DTF, rm.na = TRUE), 
  GettingElectricity = mean(Getting.Electricity...DTF, rm.na = TRUE), 
  RegisteringProperty = mean(Registering.Property...DTF, rm.na = TRUE), 
  GettingCredit = mean(Getting.Credit...DTF, rm.na = TRUE), 
  ProtectingInvestor = mean(Protecting.Minority.Investors...DTF, rm.na = TRUE), 
  PayingTaxes = mean(Paying.Taxes...DTF, rm.na = TRUE), 
  Trading = mean(Trading.across.Borders...DTF, rm.na = TRUE), 
  EnforcingContracts = mean(Enforcing.Contracts...DTF, rm.na = TRUE)) 

#frontier$Year <- gsub("DB*", "", frontier$Year) 

#frontier$Year <- as.Date(frontier$Year, "%Y")

## Theme
#install.packages("extrafont");
#library(extrafont)
#font_import(pattern = "Cambria")
#loadfonts(quiet = T)
theme_sm <- function () { 
  theme_grey(base_size=9, base_family="Avenir") %+replace% 
    theme(
      plot.background = element_blank(),
      panel.background  = element_rect(fill="gray96", colour=NA), 
      #plot.background = element_rect(fill = "lightskyblue1",colour = "pink",size = 0.5, linetype = "longdash"),
      panel.border = element_rect(fill = NA, colour = "gray60", size = 0.5), 
      panel.grid.major = element_line(colour = "grey90", size = 0.2), 
      panel.grid.minor = element_line(colour = "grey98", size = 0.5), 
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key = element_rect(fill="transparent", colour=NA),
      legend.position = "right", 
      plot.title = element_text(size=14, family = "Georgia", face = "bold", hjust = 0),
      plot.subtitle = element_text(size=10, hjust = 0),
      axis.title = (element_text(size=10)),
      plot.caption = (element_text(size = 6, hjust = 1, face = "italic"))
      #legend.key = element_rect(fill = "lightskyblue1", color = "lightskyblue1"),
      #legend.background = element_rect( fill = "lightskyblue1",color = "pink", size = 0.5,linetype = "longdash"),
      )
}

theme_sm_axis <- function() {
theme(axis.line.y = element_line(colour = "black", size = 0.5),
      axis.line.x = element_line(colour = "black", size = 0.5),
      panel.border = element_blank(),
      panel.background  = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.minor.y = element_line(colour = "grey80", size = 0.2),
      panel.grid.major.y = element_line(colour = "grey80", size = 0.2)
    )
}

### Color Palatte
#continuous_pal = c("cadetblue", "cadetblue3", "cadetblue2", "cadetblue1", "darkslategray1", "aquamarine", 'aquamarine1', 'aquamarine2', 'aquamarine3')
#discrete_pal = c('cornflowerblue','aquamarine3','deeppink3', "darkolivegreen3", "darkgoldenrod3", 'antiquewhite3', "chartreuse2", "darkslategray2", "blue")
continuous_pal = 'YlGn'
discrete_pal = 'Spectral'
qualitative_pal = 'Set2'
```

**Graph 1**
Unsurprisingly, the frontier scores are lowest for low income countries. We can see that there is a drop across frontier scores for all 4 - High Income, Upper Middle Income, Lower Middle Income and Low Income countries. We can observe that this is mostly due to Getting Electrify.
I would like to explore that further - where is the maximum drop across countries

```{r echo= FALSE, warning = FALSE, message=FALSE }
ggplot(data = melt(frontier, id.var = c("Year", "region", "incomegroup"), variable.name = "Regulatory_Process", value.name = "Score"),
       aes(x = Year, y = Score, fill = factor(incomegroup, levels = c("High income", "Upper middle income", "Lower middle income",  "Low income")))) + 
         geom_bar(stat = "identity") +
  scale_fill_brewer(palette = qualitative_pal)+
  facet_wrap(~Regulatory_Process, ncol = 3) +
  guides(fill=guide_legend(title = "Income Groups")) +
  scale_y_continuous() +
  theme_light() +
  theme_sm() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Distance to Frontier across Years \n",
       subtitle = "Average year on year improvement in distance to frontier scores \n",
       x = "Year", y = "Aggregate Frontier Scores",
       caption = 'Data from WorldBank:doingbusiness')
```

\newpage

**Graph 2 **
As we can see, the worst performing countries belong to the developing world. Among them I would like to see which ones have improved the most.

```{r, results = FALSE, echo= FALSE, warning = FALSE, message=FALSE }
# Read the shapefiles into R
country <- readOGR(dsn=path.expand("~/Documents/dataviz project/ne_110m_admin_0_countries"), layer = "ne_110m_admin_0_countries")

## SpatialPolygonsDataFrame
class(country)

# Let's transform them into tidy data that ggplot can use
country.points <- tidy(country, region = "ADM0_A3_IS")

# And join in the original variables from the shapefile
country.df <- left_join(country.points, country@data, by = c("id" = "ADM0_A3_IS"))
#join with the geospatial data
country.df <- left_join(country.df, select(doingbiz, c("ISO.alpha3.Code", "Economy", "Ease.of.Doing.Business.Rank")), by = c("id" = "ISO.alpha3.Code"))

#Map
ggplot(data=country.df, aes(long, lat, group=group, fill=Ease.of.Doing.Business.Rank)) + 
  geom_polygon() +
  geom_path(color="white") + 
  scale_fill_distiller(palette = "Spectral") +
  scale_x_continuous() + scale_y_continuous() + 
  theme_sm() +
  theme(panel.background  = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.minor.y = element_line(colour = "grey80", size = 0.2),
        panel.grid.major.y = element_line(colour = "grey80", size = 0.2),
        panel.grid.minor.x = element_line(colour = "grey80", size = 0.2),
        panel.grid.major.x = element_line(colour = "grey80", size = 0.2)) +
  labs(x = "Longitude", y = "Latitude", title = "Ease of Doing Business \n", 
       subtitle = "Worst Ranked countries are african, south america and south asia. \n",
       caption = 'Data from WorldBank:doingbusiness')
```

\newpage

**Graph 3**
It is very interesting to observe all 190 countries in the same graph. We can see that almost all high income countries (red in graph) have been stable in terms of overall distance to the frontier, ie not much change in terms of regulations towards ease of business. A surprising exception to this is Singapore.
It is surprising to see that many of the upper middle income countries (purple in graph) have improved in their performance towards ease of business,
The low income countries have had more fluctuations in terms of performance, with few notable improvements being - India, Kenya, Rwanda and Uganda. These are few of the countries I will explore in the next few graphs.

```{r echo= FALSE, warning = FALSE, message=FALSE }
#High Income Group
ggplot(data = filter(doingbiz, incomegroup == "High income"), aes(x = Year, y = Overall.DTF, color = incomegroup)) + 
  geom_line(stat = "identity") + facet_wrap(~Economy, ncol = 10) + 
  scale_y_continuous() + scale_x_continuous() + 
  scale_color_manual(values = "deeppink3") +
  guides(color = F) +
  theme_sm() + theme( strip.background = element_blank(), axis.text.x = element_blank()) +
  labs(x = "Overall DTF", y = "Years", title = "Change in High Income countries performance 2004-17")
#Upper Middle Income Group
ggplot(data = filter(doingbiz, incomegroup == "Upper middle income"), aes(x = Year, y = Overall.DTF, color = incomegroup)) + 
  geom_line(stat = "identity") + facet_wrap(~Economy, ncol = 10) + 
  scale_color_manual(values = "blue") +
  scale_y_continuous() + scale_x_continuous() + 
  guides(color = F) +
  theme_sm() + theme( strip.background = element_blank(), axis.text.x = element_blank()) +
  labs(x = "Overall DTF", y = "Years", title = "Change in Upper middle Income countries performance 2004-17")
# Lower Middle income Group
ggplot(data = filter(doingbiz, incomegroup == "Lower middle income"), aes(x = Year, y = Overall.DTF, color = incomegroup)) + 
  geom_line(stat = "identity") + facet_wrap(~Economy, ncol = 10) + 
  scale_color_manual(values = "orange") +
  scale_y_continuous() + scale_x_continuous() + 
  guides(color = F) +
  theme_sm() + theme( strip.background = element_blank(), axis.text.x = element_blank()) +
  labs(x = "Overall DTF", y = "Years", title = "Change in Lower middle Income countries performance 2004-17")
#Low Income Group
ggplot(data = filter(doingbiz, incomegroup == "Low income"), aes(x = Year, y = Overall.DTF, color = incomegroup)) + 
  geom_line(stat = "identity") + facet_wrap(~Economy, ncol = 10) + 
  scale_color_manual(values = "red") +
  scale_y_continuous() + scale_x_continuous() + 
  guides(color = F) +
  theme_sm() + theme( strip.background = element_blank(), axis.text.x = element_blank()) +
  labs(x = "Overall DTF", y = "Years", title = "Change in Low Income countries performance 2004-17"
       , caption = 'Data from WorldBank:doingbusiness')
```

\newpage

**Graph 4**

On further inspection of Singapore, we can see that, there has a steep drop in the performance of Singapore, especially in getting credit. we will use *getting credit* to further explore other countries too.


```{r echo= FALSE, warning = FALSE, message=FALSE }
data <- doingbiz %>% filter(Economy == "Singapore") %>% group_by(Year) %>% 
  summarise(StartingBusiness = mean(Starting.a.Business...DTF, rm.na = TRUE), 
            GettingConstructionPermits = mean(Dealing.with.Construction.Permits...DTF, rm.na = TRUE), 
            GettingElectricity = mean(Getting.Electricity...DTF, rm.na = TRUE), 
            RegisteringProperty = mean(Registering.Property...DTF, rm.na = TRUE), 
            GettingCredit = mean(Getting.Credit...DTF, rm.na = TRUE), 
            ProtectingInvestor = mean(Protecting.Minority.Investors...DTF, rm.na = TRUE), 
            PayingTaxes = mean(Paying.Taxes...DTF, rm.na = TRUE), 
            Trading = mean(Trading.across.Borders...DTF, rm.na = TRUE), 
            EnforcingContracts = mean(Enforcing.Contracts...DTF, rm.na = TRUE)) %>%
  melt(id.vars = "Year", variable.name = "Regulatory_Process", value.name = "Score")

ggplot() + 
  geom_line(data = filter(data, (Regulatory_Process != "Trading" & Regulatory_Process != "GettingCredit" & Regulatory_Process != "ProtectingInvestor")), aes(x = Year, y = Score, group = Regulatory_Process), color = 'gray86') + 
  geom_line(data = filter(data, (Regulatory_Process == "Trading" | Regulatory_Process == "GettingCredit" | Regulatory_Process == "ProtectingInvestor")), aes(x = Year, y = Score, color = Regulatory_Process)) + 
  scale_x_discrete(expand=c(0, 1)) +
  geom_dl(data = filter(data, (Regulatory_Process == "Trading" | Regulatory_Process == "GettingCredit" | Regulatory_Process == "ProtectingInvestor")), aes(x = Year, y = Score,label = Regulatory_Process), method = list(dl.trans(x = x - 0.2), dl.combine("first.points"), cex = 0.6)) +
  guides(color = F) +
  scale_color_brewer(palette = qualitative_pal) + 
  theme_sm() + theme(plot.background = element_blank(),
      panel.background  = element_blank(),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(),
      panel.grid.minor.x = element_line(colour = "grey80", size = 0.2),
      panel.grid.major.x = element_line(colour = "grey80", size = 0.2)) +
  labs(x = "Year of Reference", y = "Scores Across different Regulatory Measures", title = "Performace drop by Singapore 2004-17 \n", 
       subtitle = "There has been a steep drop in Sgp's performance in the financial aspects. However, it is still ranked 7 overall \n", 
       caption = 'Data from WorldBank:doingbusiness')
rm(data)
```
\newpage

**Graph 5**

*Dealing with Construction Permits* It is a measure of procedures involved, time taken, cost of construction and building quality control index. India is the worst performer in this category. Though it is trying to catch up. Other countries are more stable in their performance.

*Getting Electricity* It is a measure of procedures involved, time taken, cost of construction and reliability of supply and transparency of tariff index. Countries have been stable in their performance except for Rwanda, who's performance dropped by 15 points.

*Registering Property* It is a measure of procedures involved, time taken, cost of construction and quality of the land administration index including equal access to property rights. Countries have been stable in their performance.

*Enforcing Contracts* It is a measure of procedures involved, time taken, cost of construction and quality of judicial processes index including weight of woman's testimony in court. Surprisingly, India is one of the worst performing countries in this aspect. Even in 2017, it is only at 35 points (with 56 being lowest among the other countries).

```{r echo= FALSE, warning = FALSE, message=FALSE }
data <- doingbiz %>% filter(Economy == "Singapore" | Economy == "India" | Economy == "Rwanda" | Economy == "Uganda") %>% 
  filter(Year == 2014 | Year == 2017) %>%
  group_by(Year, Economy) %>% 
  summarise(StartingBusiness = mean(Starting.a.Business...DTF, rm.na = TRUE), 
            GettingConstructionPermits = mean(Dealing.with.Construction.Permits...DTF, rm.na = TRUE), 
            GettingElectricity = mean(Getting.Electricity...DTF, rm.na = TRUE), 
            RegisteringProperty = mean(Registering.Property...DTF, rm.na = TRUE), 
            GettingCredit = mean(Getting.Credit...DTF, rm.na = TRUE), 
            ProtectingInvestor = mean(Protecting.Minority.Investors...DTF, rm.na = TRUE), 
            PayingTaxes = mean(Paying.Taxes...DTF, rm.na = TRUE), 
            Trading = mean(Trading.across.Borders...DTF, rm.na = TRUE), 
            EnforcingContracts = mean(Enforcing.Contracts...DTF, rm.na = TRUE)) %>%
  melt(id.var = c("Year", 'Economy'), variable.name = "Regulatory_Process", 
       value.name = "Score")
data$Year <- as.character(data$Year)
ggplot(filter(data, Regulatory_Process == 'GettingConstructionPermits'| Regulatory_Process == 'GettingElectricity'|
                Regulatory_Process == 'RegisteringProperty' | Regulatory_Process == 'EnforcingContracts'),
       aes(x = Year, y = log(Score))) +
  guides(color = guide_legend("Country")) +
  geom_point(aes(color = Economy), size = 2)+
  geom_line(aes(x = Year, y = log(Score), group = Economy, color = Economy)) +
  facet_wrap(~Regulatory_Process, ncol = 2) +
  scale_fill_brewer(palette = discrete_pal)+
  geom_text(aes(x = Year, label = Score, y = log(Score)),
            size = 2,  hjust = 0, check_overlap = TRUE) +
  theme_sm() + theme(legend.position = "bottom") + theme_sm_axis() + 
  labs(x = "", y = "Performance of the country",
       title = "Comparative performance between 2014 and 2016 \n",
       caption = 'Data from WorldBank:doingbusiness')
#Next 4 features
ggplot(filter(data, Regulatory_Process != 'GettingConstructionPermits'& Regulatory_Process != 'GettingElectricity'&
                Regulatory_Process != 'StartingBusiness' & Regulatory_Process != 'RegisteringProperty' & Regulatory_Process != 'EnforcingContracts'),
       aes(x = Year, y = log(Score))) +
  guides(color = guide_legend("Country")) +
  geom_point(aes(color = Economy), size = 2)+
  geom_line(aes(x = Year, y = log(Score), group = Economy, color = Economy)) +
  facet_wrap(~Regulatory_Process, ncol = 2) +
  scale_fill_brewer(palette = discrete_pal)+
  geom_text(aes(x = Year, label = Score, y = log(Score)),
            size = 3,  hjust = 0, check_overlap = TRUE) +
  theme_sm() + theme(legend.position = "bottom") + theme_sm_axis() + 
  labs(x = "Year of Interest", y = "Performance of the country",
       caption = 'Data from WorldBank:doingbusiness')
```

*Getting Credit* Getting Credit is a measure of the degree to which collateral and bankruptcy laws protect the rights of borrowers and lenders and depth of credit information index of a country. Uganda's has quickly risen to catch up with the other countries. It moved from a score of 43 to 65. India and Rwanda have been stable. Singapore's performance dropped by 20 points too.

*Protecting Investors*  It is a measure of extent of conflict of interest regulation (protection of shareholders against directors’ misuse of corporate assets for personal gain) and extent of shareholder  in corporate governance. Rwanda has become worse in terms of protecting investors.

*Paying Taxes* The tax payments indicator reflects the total number of taxes and contributions paid, the method of payment, the frequency of payment, the frequency of filing and the number of agencies involved; time taken to file taxes; total tax rate; and post filing procedures. India is the only country with significant drop in the performance. This is surprising as there have been stricter rules being implemented there.

*Trading* Trading across Borders is a measure of documents needed for export; time to export; cost to export; documents needed for import; time taken for import; and costs involved in import.
Both India and Singapore's performance has deteriorated, whereas Uganda and Rwanda's performance has improved. 


\newpage
**Graph 6**

From this graph we can see that though the performance of Uganda and Rwanda has improved, whereas Singapore and India has worsened. Surprisingly,  Rwanda's final performance is better than that of Sgp (current rank 7). 
I would like to explore this trend of increase in *ease of getting credit* in the two sub Saharan countries to the other sub Saharan countries

```{r echo= FALSE, warning = FALSE, message=FALSE }
data <- doingbiz %>% filter(Economy == "Singapore" | Economy == "India" | Economy == "Rwanda" | Economy == "Uganda") %>% 
  filter(Year == 2014 | Year == 2017) %>%
  group_by(Year, Economy) %>% 
  summarise(StartingBusiness = mean(Starting.a.Business...DTF, rm.na = TRUE), 
            GettingConstructionPermits = mean(Dealing.with.Construction.Permits...DTF, rm.na = TRUE), 
            GettingElectricity = mean(Getting.Electricity...DTF, rm.na = TRUE), 
            RegisteringProperty = mean(Registering.Property...DTF, rm.na = TRUE), 
            GettingCredit = mean(Getting.Credit...DTF, rm.na = TRUE), 
            ProtectingInvestor = mean(Protecting.Minority.Investors...DTF, rm.na = TRUE), 
            PayingTaxes = mean(Paying.Taxes...DTF, rm.na = TRUE), 
            Trading = mean(Trading.across.Borders...DTF, rm.na = TRUE), 
            EnforcingContracts = mean(Enforcing.Contracts...DTF, rm.na = TRUE)) 
data$Year <- as.character(data$Year)
h <- mean(doingbiz$Getting.Credit...DTF, na.rm = TRUE) %>% round(digit = 1)
ggplot(data,aes(x = Economy, y = GettingCredit, fill = Year)) + 
  scale_fill_brewer(palette = continuous_pal)+
  geom_bar(stat = "identity",position = "dodge" ) + 
  geom_hline(aes(yintercept = h)) +
  annotate(geom="text", label= h, x=0, y=h, vjust=-1) +
  theme_sm() + theme_sm_axis() +
  labs(x = "Countries of Interest", y = "Ease of getting credit in the country",
     title = "Comparative performance between 2014 and 2016 \n", 
     subtitle = "All countries of interest perform better than the average of all countries \n",
     caption = 'Data from WorldBank:doingbusiness')
```

\newpage


**Part 2:**

We are now going to focus on one country that performed even better than Singapore - Rwanda. 
I will be using Enterprise Survey data for this part. Enterprise Survey focuses on many factors that shape a business environment. The Enterprise Survey is conducted by World Bank and its partners across the world on all small, medium and large firms.

**Graph 6: Performance of Rwanada vs other Sub Saharan Countries**
As we can see in the graph, the electricity condition, corruption and political instability is not one of the major concern of Rwanda's enterprises, unlike the rest of the Saharan countries. However, when in comes to taxes and regulation, Rwanda is lagging behind. I will be exploring it further in future graphs.


```{r echo= FALSE, warning = FALSE, message=FALSE }
#Graph 6 (not functional) : Rwanda's Ranking of Obstacles in terms of other Sub Saharan Countries
entrep_data <- read_csv("obstacle.csv")
cols.num <- c('Access to finance',	'Access to land',	'Business licensing and permits',	'Corruption',	'Courts',	'Crime, theft and disorder',	'Customs and trade regulations',	'Electricity',	'Inadequately educated workforce',	'Labor regulations', 'Political instability',	'Practices of the informal sector',	'Tax administration',	'Tax rates', 'Transportation')
entrep_data[cols.num] <- sapply(entrep_data[cols.num],as.numeric)
#filtering data to contain only current data
rwanda <- filter(entrep_data, (Economy == "Rwanda" | Economy == "Sub-Saharan Africa") & (Year == 2011 | is.na(Year)) & is.na(Subgroup)) %>%
  select(-(Year:`Average/SE/N`), -`Business licensing and permits`, -Courts, - `Crime, theft and disorder`, - `Inadequately educated workforce`, -`Labor regulations`) %>%
  melt(id.var = c("Economy"), variable.name = "Obstacle", value.name = "Percentage")

levels(rwanda$Obstacle) <- gsub(" ", "\n", levels(rwanda$Obstacle))
ggplot() + 
  geom_bar(data = filter(rwanda, Economy == "Rwanda"), aes(x = Obstacle, y = Percentage), stat = "Identity", fill = "cornflowerblue") +
  geom_point(data = filter(rwanda, Economy == "Sub-Saharan Africa"), aes(x = Obstacle, y = Percentage), fill = "orange", shape = 23) +
  guides(fill = guide_legend(title = "Sub-Saharan Africa")) +
  theme_sm() + theme_sm_axis() + theme(legend.position = "top") +
  labs(x = "Obstacles faced by firms", y = "Percent of Firms", title = "Rwanda's percentage of firms that identify the problem \n as the biggest obstacle, 2011 \n",
         subtitle = "Top obstacles identified by Rwanda's firms benchmarked against the regional average (sub saharan africa)", 
         caption = 'Data from http://www.enterprisesurveys.org')
```

\newpage

**Graph 7**
We see a remarkable improvement in electricity (reduction in outages and increase in number of firms with generators). Political instability has almost stopped being an issue. However, as ease of doing business has improved in Rwanda, there has been an increase in concern related to the informal business sector. Taxes and regulations continue to remain a major concern for the firms.

```{r echo= FALSE, warning = FALSE, message=FALSE }
#Graph 7 : Rwanda's Improvement
entrep_data <- read_csv("obstacle.csv")
rwanda <- filter(entrep_data, Economy == "Rwanda" & `Subgroup Level` == "All")
cols.num <- c('Access to finance',	'Access to land',	'Business licensing and permits',	'Corruption',	'Courts',	'Crime, theft and disorder',	'Customs and trade regulations',	'Electricity',	'Inadequately educated workforce',	'Labor regulations', 'Political instability',	'Practices of the informal sector',	'Tax administration',	'Tax rates', 'Transportation')
rwanda[cols.num] <- sapply(rwanda[cols.num],as.numeric)
rwanda <- mutate(rwanda, Other = `Business licensing and permits` +	`Corruption` + `Courts` +	`Crime, theft and disorder` + `Customs and trade regulations` + `Labor regulations` + `Inadequately educated workforce`,
                 Tax  = `Tax administration` + `Tax rates`) %>%
  select(-Subgroup, -`Subgroup Level`, -`Average/SE/N`, -Economy, -(`Business licensing and permits`: `Customs and trade regulations`), -`Labor regulations`, -`Tax rates`, - `Tax administration`, -`Inadequately educated workforce`) %>%
  melt(id.var = c("Year", "Top Subgroup Level"), variable.name = "Obstacle", value.name = "Percentage")

ggplot(data=filter(rwanda) , aes(x = "", y = Percentage, fill = Obstacle)) +
  facet_grid(Year ~`Top Subgroup Level`) +
  geom_col(width = 1) +
  scale_fill_brewer(name = "Percent of firms that believe following \n to be the biggest obstacle", palette = qualitative_pal)+
  coord_polar("y") +
  theme_sm() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(x = "", y = "", title = "Rwanda's Profile in terms of Business Environment \n Obstacles (2006 vs 2011) \n",
subtitle = "In 2006 20 firms were surveyed in Butare (9 in 2011) and 192 in Kigali (232 in 2011). Inacessibility 
to electricity has stopped being an issue. However, an increase in practices of the informal 
sector seems to be a rising concern. Tax and Regulation is still a big concern.  \n", 
caption = 'Data from http://www.enterprisesurveys.org')

```
\newpage

**Graph8**
An average firm in Rwanda has more number of female employees/owners as compared to the rest of sub Saharan Africa. Is this because of improvement in political stability or reduction in corruption or improvement in finances and technology. I would like to explore the relationship between these in future. 
As we can see, most of the firms are privately owned (domestic and foreign). Does this result in more inequity? 

```{r echo= FALSE, warning = FALSE, message=FALSE }
female <- read_csv("gender.csv") %>% 
  filter((Economy == "Rwanda" | Economy == "Sub-Saharan Africa") & (Year == 2011 | is.na(Year)) & is.na(Subgroup)) %>%
  select(Economy, `Proportion of permanent full-time workers that are female (%)`, `Percent of firms with female participation in ownership`)
ownership <- read_csv("firm.csv") %>%
  filter((Economy == "Rwanda" | Economy == "Sub-Saharan Africa") & (Year == 2011 | is.na(Year)) & is.na(Subgroup)) %>%
  select(Economy, `Proportion of private domestic ownership in a firm (%)`, `Percent of firms with at least 10% of foreign ownership`, `Percent of firms with at least 10% of government/state ownership`,`Percent of firms with an internationally-recognized quality certification`)
firms <- merge(female, ownership, by = "Economy") %>%
  melt(id.var = c('Economy'), variable.name = "Charcteristic", value.name = "Percentage")
firms$Percentage <- as.numeric(firms$Percentage)

levels(firms$Charcteristic) <- gsub(" ", "\n", levels(firms$Charcteristic))
ggplot(firms, aes(x = Charcteristic, y = Percentage, fill = Economy)) + 
       geom_bar(position = 'dodge', stat = 'identity')+
  scale_fill_brewer(palette = qualitative_pal) +
  theme_sm() +
  labs(x = "", y = "Percentage of Firms", title = "Rwanda's Country Profile (2011) \n",
       subtitle = "It is interesting to observe that % ownership by women is higher than the working full time \n female population. Most of the firms are private run as compared to government participation.\n", 
       caption = 'Data from http://www.enterprisesurveys.org')
```

