---
title: 'Comparing Business Regulation for Domestic Firms in 190 Economies: Equal Opportunity
  for All?'
author: "Shambhavi Mohan"
date: 'Due: October 15, 2017'
output:
  pdf_document: default
  html_document: default
urlcolor: green
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Background
The opportunity to find a job or ease of doing business can help people come out of poverty or prevent them from going in. I want to explore how the ease of doing business has improved across economies and how different lower income economies are as compared to high income ones.

*Important Definition:* An economy’s distance to frontier is reflected on a scale from 0 to 100, where 0 represents the lowest performance and 100 represents the frontier. For example, a score of 75 in DB 2016 means an economy was 25 percentage points away from the frontier constructed from the best performances across all economies and across time.


```{r, echo=FALSE, warning=FALSE, message=FALSE}

setwd("~/Documents/dataviz project")

#import libraries
library(readr)
library(haven)
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(reshape2)
library(rworldmap)
library(RColorBrewer)
library(extrafont)
library(directlabels)


doingbiz <- read.csv("doingbusiness.csv")

#Changing names for Economy
doingbiz$Economy <- str_replace(doingbiz$Economy, "Russian Federation", "Russia" )
doingbiz$Economy <- str_replace(doingbiz$Economy, "Antigua and Barbuda",  "Antigua")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Brunei Darussalam", "Brunei")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Cabo Verde", "Cape Verde")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Congo, Dem Rep" , "Democratic Republic of the Congo")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Congo, Rep", "Republic of Congo")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Cote d’Ivoire", "Ivory Coast")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Egypt, Arab Rep", "Egypt")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Gambia, The", "Gambia")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Iran, Islamic Rep", 'Iran')
doingbiz$Economy <- str_replace(doingbiz$Economy, "Korea, Rep", "South Korea")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Kyrgyz Republic", "Kyrgyzstan")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Lao PDR", "Laos")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Macedonia, FYR", "Macedonia")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Micronesia, Fed Sts", "Micronesia")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Puerto Rico US", "Puerto Rico")
doingbiz$Economy <- str_replace(doingbiz$Economy, "São Tomé and Príncipe", "Sao Tome and Principe")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Slovak Republic", "Slovakia")
doingbiz$Economy <- str_replace(doingbiz$Economy, "St Kitts and Nevis", "Saint Kitts")
doingbiz$Economy <- str_replace(doingbiz$Economy, "St Lucia", "Saint Lucia")
doingbiz$Economy <- str_replace(doingbiz$Economy, "St Vincent and the Grenadines", "Saint Vincent")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Syrian Arab Republic", "Syria")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Taiwan, China", "Taiwan")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Trinidad and Tobago", "Trinidad" )
doingbiz$Economy <- str_replace(doingbiz$Economy, "United Kingdom", "UK")
doingbiz$Economy <- str_replace(doingbiz$Economy, "United States", "USA")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Venezuela, RB", "Venezuela")
doingbiz$Economy <- str_replace(doingbiz$Economy, "West Bank and Gaza", "Palestine")
doingbiz$Economy <- str_replace(doingbiz$Economy, "Yemen, Rep", "Yemen")

doingbiz$Year <- gsub("DB*", "", doingbiz$Year) 
doingbiz$Year <- as.numeric(doingbiz$Year)
doingbiz <- doingbiz %>% filter(!str_detect(Economy, "-"))

#for the world map
map.world <- map_data(map="world")
map.world$region <- str_replace(map.world$region, "Tobago", "Trinidad")
map.world$region <- str_replace(map.world$region, "Virgin Island", "USA")

names(doingbiz)[names(doingbiz) == "region"] <- 'area' #toavoid conflict later

#Only distance to frontier data
frontier <- doingbiz %>% group_by(area,incomegroup, Year) %>% 
  summarise(StartingBusiness = mean(Starting.a.Business...DTF, rm.na = TRUE), 
  GettingConstructionPermits = mean(Dealing.with.Construction.Permits...DTF, rm.na = TRUE), 
  GettingElectricity = mean(Getting.Electricity...DTF, rm.na = TRUE), 
  RegisteringProperty = mean(Registering.Property...DTF, rm.na = TRUE), 
  GettingCredit = mean(Getting.Credit...DTF, rm.na = TRUE), 
  ProtectingInvestor = mean(Protecting.Minority.Investors...DTF, rm.na = TRUE), 
  PayingTaxes = mean(Paying.Taxes...DTF, rm.na = TRUE), 
  Trading = mean(Trading.across.Borders...DTF, rm.na = TRUE), 
  EnforcingContracts = mean(Enforcing.Contracts...DTF, rm.na = TRUE)) 
```

**Graph 1**
Unsurprisingly, the frontier scores are lowesst for low income countries. We can see that there is a drop across frontier scores for all 4 - High Income, Upper Middle Income, Lower Middle Income and Low Income countries. We can observe that this is mostly due to Getting Electrity.
I would like to explore that further - where is the maximum drop across countries

```{r echo= FALSE, warning = FALSE, message=FALSE }
#Graph1
ggplot(data = melt(frontier, id.var = c("Year", "area", "incomegroup"), variable.name = "Regulatory_Process", value.name = "Score"),
       aes(x = Year, y = Score, fill = Regulatory_Process)) + geom_bar(stat = "identity") +
  scale_color_brewer(type = "seq") +
  facet_wrap(~incomegroup, ncol = 2) +
  scale_y_continuous() +
  theme_light() +
  theme(panel.background = element_rect(fill = "azure"),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Distance to Frontier across Years",
       subtitle = "Average year on year improvement in distance to frontier scores",
       x = "Year", y = "Aggregate Frontier Scores",
       caption = 'Data from WorldBank:doingbusiness')
```

\newpage

**Graph 2 **
As we can see, the worst performing countries belong to the developing world. Amongst them I would like to see which ones have improved the most.

```{r echo= FALSE, warning = FALSE, message=FALSE }
#data is merged dataset of our data:doingbiz and worldmap data
ggplot() + 
  geom_polygon(data=merge(x = map.world, y = filter(doingbiz, Year == 2016), by.x = 'region', by.y = "Economy"), aes(x=long, y=lat, group = group, fill=Ease.of.Doing.Business.Rank)) + 
  geom_jitter()+ 
  scale_fill_continuous(low = "green", high = "red", guide="colorbar", guide_legend("Country Rankings") ) + 
  scale_x_continuous() + scale_y_continuous() + 
  labs(x = "Longitude", y = "Latitude", title = "Ease of Doing Business, 2016", subtitle = "Worst Ranked countries are African, South Americas and South Asian", caption = 'Data from WorldBank:doingbusiness')

```

\newpage

**Graph 3**
It is very interesting to observe all 190 countries in the same graph. We can see that almost all high income countries (red in graph) have been stable in terms of overall distance to the frontier, ie not much change in terms of regulations towards ease of business. A surprising exception to this is Singapore.
It is surprising to see that many of the upper middle income countries (purple in graph) have improved in their performace towards ease of business,
The low income countries have had more fluctuations in terms of performancem, with few notable imrovements being - India, Kenya, Rwanda and Uganda. These are few of the countries I will explore in the next few graphs.

```{r echo= FALSE, warning = FALSE, message=FALSE }
  ggplot(data = doingbiz, aes(x = Year, y = Overall.DTF, color = incomegroup)) +
  geom_line(stat = "identity") + facet_wrap(~Economy, ncol = 15) + 
  scale_y_continuous() + scale_x_continuous() + 
  geom_dl(aes(label = Economy), method = list(dl.combine("first.points"), rot = 45, cex = 0.5)) +
  guides(color = F) +
  theme(strip.text.y = element_blank(), strip.text.x = element_blank(), strip.background = element_blank(), strip.text = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank()) +
  labs(y = "Overall DTF", x = "Years", title = "Change in countries perfomances 2004-17", 
       subtitle = "Here colors represent the income group, with red for the highest income and green for the least income", caption = 'Data from WorldBank:doingbusiness')

```

\newpage

**Graph 4**

On further inspection of Singapore, we can see that, there has a steep drop in the performance of Singapore, especially in getting credit. we will use *getting credit* to further explore other countries too.


```{r echo= FALSE, warning = FALSE, message=FALSE }
data <- doingbiz %>% filter(Economy == "Singapore") %>% group_by(Year) %>% 
  summarise(StartingBusiness = mean(Starting.a.Business...DTF, rm.na = TRUE), 
            GettingConstructionPermits = mean(Dealing.with.Construction.Permits...DTF, rm.na = TRUE), 
            GettingElectricity = mean(Getting.Electricity...DTF, rm.na = TRUE), 
            RegisteringProperty = mean(Registering.Property...DTF, rm.na = TRUE), 
            GettingCredit = mean(Getting.Credit...DTF, rm.na = TRUE), 
            ProtectingInvestor = mean(Protecting.Minority.Investors...DTF, rm.na = TRUE), 
            PayingTaxes = mean(Paying.Taxes...DTF, rm.na = TRUE), 
            Trading = mean(Trading.across.Borders...DTF, rm.na = TRUE), 
            EnforcingContracts = mean(Enforcing.Contracts...DTF, rm.na = TRUE)) 

ggplot(data = melt(data, id.vars = "Year", variable.name = "Regulatory_Process", value.name = "Score"), aes(x = Year, y = Score, color = Regulatory_Process)) + 
  geom_line() + scale_color_brewer(palette = "Set1") + 
  labs(x = "Year", y = "Scores Across different Regulatory Measures", title = "Performace drop by Singapore 2004-17", 
       subtitle = "There has been a steep drop in Sgp's performance. However, it is still ranked 7 overall", caption = 'Data from WorldBank:doingbusiness')
rm(data)
```
\newpage

**Graph 5**

From this graph we can see that though the performace of Uganda and Rwanda has improved, whereas Singapore and India has worsened. Surprisingly,  Rwanda's final performance is better than that of Sgp (current rank 7). 
I would like to explore this trend of increase in *ease of getting credit* in the two sub saharan countries to the other sub saharan countries

```{r echo= FALSE, warning = FALSE, message=FALSE }
data <- doingbiz %>% filter(Economy == "Singapore" | Economy == "India" | Economy == "Rwanda" | Economy == "Uganda") %>% 
  filter(Year == 2014 | Year == 2017) %>%
  group_by(Year, Economy) %>% 
  summarise(StartingBusiness = mean(Starting.a.Business...DTF, rm.na = TRUE), 
            GettingConstructionPermits = mean(Dealing.with.Construction.Permits...DTF, rm.na = TRUE), 
            GettingElectricity = mean(Getting.Electricity...DTF, rm.na = TRUE), 
            RegisteringProperty = mean(Registering.Property...DTF, rm.na = TRUE), 
            GettingCredit = mean(Getting.Credit...DTF, rm.na = TRUE), 
            ProtectingInvestor = mean(Protecting.Minority.Investors...DTF, rm.na = TRUE), 
            PayingTaxes = mean(Paying.Taxes...DTF, rm.na = TRUE), 
            Trading = mean(Trading.across.Borders...DTF, rm.na = TRUE), 
            EnforcingContracts = mean(Enforcing.Contracts...DTF, rm.na = TRUE)) 

data$Year <- as.character(data$Year)

h <- mean(doingbiz$Getting.Credit...DTF, na.rm = TRUE) %>% round(digit = 1)

ggplot(data,aes(x = Economy, y = GettingCredit, fill = Year)) + 
  geom_bar(stat = "identity",position = "dodge" ) + 
  geom_hline(aes(yintercept = h)) +
  annotate(geom="text", label= h, x=1, y=h, vjust=-1) +
  labs(x = "Countries of Interest", y = "Ease of getting credit in the country",
     title = "Comparative performance between 2014 and 2016", 
     subtitle = "All countries of interest perform better than the mean")
```